---
title: "module_two_communication_ahaworth"
author: "Andrew Haworth"
date: today
format:
  html: default
  pdf: default
  docx: default
bibliography: references.bib
keywords:
- climate change
- carbon dixide
abstract: "This resport examines changes in gloabl warming"
params:
  station: "Mauna Loa"
  start_year: 1990
---

Start to write text.

- Global temperature anomaly (NASA GISTEMP)
- Atmospheric CO~2~ concentration (Mauna Loa Observatory)
- [Data source](https://climate.nasa.gov/vital-signs/)

> CO2 concentration has risen every year on record since systematic measurement began.

## Data
Here are my data files

## Results

## Code chunks and inline code
```{r}
data_path <- "https://websites.umich.edu/~zhukai/env-data-sci/data"
mauna_loa_co2 <- read.table(file.path(data_path, "climate", "mauna_loa_co2.txt"), skip = 42, header = FALSE, na.strings = c("-1", "-9.99", "-0.99"))
colnames(mauna_loa_co2) <- c("year", "month", "decimal_date", "average", "interpolated", "n_day", "stdv_day", "uncertainty_month")
```
## Add the parameter R code

```{r}
report_station <- params$station
report_start <- params$start_year

subset(mauna_loa_co2, year >= report_start)
```

## Other options target specific kinds of output rather than the whole chunk: message: false and warning: false suppress messages and warnings, results: hide and fig-show: hide hide printed output and plots, and error: true lets rendering continue past a failing chunk instead of stopping the whole document.
```{r}
#| message: false
#| warning: false

# Summarize CO2 by decade for the report
mauna_loa_co2$decade <- floor(mauna_loa_co2$year / 10) * 10
aggregate(interpolated ~ decade, data = mauna_loa_co2, FUN = mean, na.rm = TRUE)
```

## Rather than repeating the same options on every chunk, we can set them once in the YAML header under execute to apply throughout the entire document.

```{r}
n_observations <- sum(!is.na(mauna_loa_co2$average))
```
There are `r n_observations` non-missing monthly CO2 observations.

```{r}
#| message: false

library(tidyverse)
```

## Figures, tables and citations 9/16/2026
```{r}
mauna_loa_co2$Date <- as.Date(with(mauna_loa_co2, ISOdate(year, month, 1)))
```

```{r}
#| label: fig-co2-trend
#| fig-width: 6
#| fig-height: 4
#| fig-cap: "Monthly atmospheric CO2 concentration measured at Mauna Loa Observatory"
#| fig-alt: "Line plot of CO2 concentration in parts per million against date, showing a steady long-term increase with a seasonal sawtooth pattern"
#| out-width: "70%"

ggplot(mauna_loa_co2, aes(x = Date, y = interpolated)) +
  geom_line(color = "darkgreen") +
  labs(x = "Date", y = "CO2 concentration (ppm)", title = "Atmospheric CO2 concentration") +
  theme_minimal()
```
As shown in @fig-co2-trend, atmospheric CO2 has risen steadily since 1958.

The complete source for the labeled table is:
```{r}
#| label: tbl-co2-decade
#| tbl-cap: "Atmospheric CO2 concentration by decade"

mauna_loa_co2$decade <- floor(mauna_loa_co2$year / 10) * 10
co2_by_decade <- aggregate(interpolated ~ decade, data = mauna_loa_co2, FUN = mean, na.rm = TRUE)
co2_by_decade$interpolated <- round(co2_by_decade$interpolated, 1)
names(co2_by_decade) <- c("Decade", "Mean CO2 (ppm)")
knitr::kable(co2_by_decade)
```

## Quarto also supports citations to external sources.

Atmospheric CO2 has been measured continuously at Mauna Loa since 1958, extending an ice-core record that reaches back hundreds of thousands of years [@barnola2003].
